---
title: "Recorder feedback delivery pipeline"
output: html_document
date: "`r Sys.Date()`"
resource_files:
  - templates_html/
  - templates/
  - R/
  - config.yml
  - data/
  - computations/
  - renders/
  - _targets.R
---

## Overview

This is an R markdown file for the full pipeline from data gathering to dispatching feedback. If you fork this repo then this should be adapted for your use case. Being an .Rmd file also means it can be published to posit connect and run using their scheduling service, hence why resource_files are included in the page yml. 

Use the code in this chunk to generate the .R version of this pipeline which you can then run from command line using `Rscript run_pipeline.R`.

```{r, eval = F}
#run this line 
knitr::purl(input = "run_pipeline.Rmd", output = "run_pipeline.R",documentation = 0)
```

## Pipeline

### Set up

The batch_id is used as an identifier for the collection of feedback items that is being generated per pipeline run. It comprises of the date and some random characters. We use `{config}` R package to load config from `config.yml`. You may also wish to source `run_preflight_tests.R` which checks connections to API etc.

```{r configuration}
#PREP
#generating batch_id
batch_id <- paste0(Sys.Date(),"-",paste(sample(letters,10),collapse = ""))

#load packages
library(httr)
library(jsonlite)
library(curl)

#load configuration from config.yml using config package
config <- config::get()
```

### Gathering

This code chunk gathers users from controller app (if you have configured it to do so), or will use the user data file you have put in `data/`. This might be a list of users you are manually maintaining or an export from a sign-up list or other email list management service.

```{r gather}
# Get users
if(config$gather_from_controller_app){
  # GATHER
  source("R/gather/get_subscribers_from_controller.R")
  # Get subscribers for the list and write the data frame
  subscribers_df <- get_subscribers_from_controller(api_url = config$controller_app_base_url, 
                                                    email_list_id = config$controller_app_list_id, 
                                                    api_token = config$controller_app_api_key) 
  write.csv(subscribers_df,config$participant_data_file,row.names = F)
}

```

This code chunks runs the script you have defined in `config$gather_bio_script`. This is where you might have written an R script specific to your use-case that connects to and downloads data from your records database for the users provided. It does a check to see if the biodiversity data file has been updates, and errors if it has now been updated.

```{r}
# get their records using this script which will save data to config$data_file
source(config$gather_bio_script)

# check that records have been updated
if(difftime(file.info(config$data_file)$mtime,Sys.time(),units = "secs") < 30){
  print("Biodiversity data file has been updated")
} else {
  stop("Biodiversity data file has not been updated, check for issue in gather script")
}
```

### Generate

This triggers the `{targets}` pipeline defined in `_targets.R`.

```{r generate}
# GENERATE
#run the pipeline
Sys.setenv(BATCH_ID = batch_id) # Set an environment variable in R for the batch code
targets::tar_make() # run the pipeline
Sys.unsetenv("BATCH_ID") #and unset the variable
```

### Send

This code chunk dispatches the feedback that has been generated by the targets pipeline. Currently this is done by email via SMTP, but could be adapted to send email via the API of an email services (e.g. sendgrid, hubspot, mailchimp).

The email send code is wrapped in a try-catch so errors can be handled.

```{r send}
#SEND

#maintain a dataframe with the email and any errors that arise when sending
status_log <- data.frame(
  user_id = character(),
  email = character(),
  status = character(),
  message = character(),
  stringsAsFactors = FALSE
)

source("R/send/send_notify_app.R")
library(blastula)

if(config$mail_creds == "envvar"){
  Sys.setenv(SMTP_PASSWORD = config$mail_password)
  creds <- creds_envvar(
    user = config$mail_username,
    pass_envvar = "SMTP_PASSWORD",
    host = config$mail_server,
    port = config$mail_port,
    use_ssl = config$mail_use_ssl)
}

if(config$mail_creds == "anonymous"){
  creds <- creds_anonymous(host = config$mail_server,port=config$mail_port,use_ssl = config$mail_use_ssl)
}

meta_table <-read.csv(paste0("renders/",batch_id,"/meta_table.csv"))
participants <- read.csv(config$participant_data_file)

#here we are going through all the users and sending email
if (!is.data.frame(meta_table)) {
  stop("meta_table must be a data frame.")
}

for (i in 1:nrow(meta_table)) {
  
  result <- tryCatch({
    # Debug: Print current row being processed
    print(paste("Processing row:", i))
    
    # Check for required columns in meta_table
    if (!all(c("content_key", "user_id", "file", "email") %in% colnames(meta_table))) {
      stop("meta_table is missing required columns.")
    }
    
    # Safely extract values
    content_key <- meta_table[i, "content_key"]
    user_id <- meta_table[i, "user_id"]
    file <- meta_table[i, "file"]
    email <- meta_table[i, "email"]
    
    # Perform operations
    if(config$gather_from_controller_app){
      send_notify_app(
        content_key = content_key,
        user_external_key = user_id,
        batch_id = batch_id,
        config = config
      )
    }
    
    sender <- config$mail_default_sender
    names(sender) <- config$mail_default_name
    email_obj <- blastula:::cid_images(file)
    
    smtp_send(
        email_obj,
        from = sender,
        to = email,
        subject = config$mail_default_subject,
        credentials = creds,
        verbose = FALSE
      )
      data.frame(
        user_id = user_id,
        email = email,
        status = "Success",
        message = "Email sent",
        stringsAsFactors = FALSE
      )
    }, error = function(e) {
      print("Email not sent")
      data.frame(
        user_id = user_id,
        email = email,
        status = "Failed",
        message = e$message,
        stringsAsFactors = FALSE
      )
    })
  status_log <- rbind(status_log,result)
}
```

This Send email to admin with status of the dispatch, how many emails were sent successfully, how many failed.

```{r}
status_lines <- apply(status_log, 1, function(row) {
  paste(
    "User ID:", row["user_id"], 
    "| Email:", row["email"], 
    "| Status:", row["status"], 
    "| Message:", row["message"]
  )
})

report_email <- compose_email(
  body = md(
    paste0(
      "### Email Send Summary",
      "
      Total attempted: ", nrow(status_log),
      "
      ✅ Success: ", sum(status_log$status == "Success"), 
      "
      ❌ Failed: ", sum(status_log$status == "Failed"),
      "
      Please check logs for more information"
    )
    ),
  footer = Sys.time()
  )

smtp_send(
  email = report_email,
  from = sender,
  to = config$mail_test_recipient,  # your email
  subject = paste("Email Batch Status Report -", Sys.Date()),
  credentials = creds,
  verbose = FALSE
)
```
