---
title: "Your Wildlife Recording Challenge!"
params:
    user_name: ""
    user_data: ""
    bg_data: ""
    bg_computed_objects: ""
    user_computed_objects: ""
    user_email: "UNKNOWN"
    content_key: ""
    config: ""
footer-date-time: "`r  format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`"
user-email: "`r params$user_email`"
---

```{css, echo=FALSE}
body{
  font-family: Helvetica;
  font-size: 13pt;
}

```

```{r, include=FALSE}
taxon <- "butterflies"
taxon_target <- 4
group <- 4
```

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(htmltools)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r calculations, include=FALSE}
max_list_length_df <- data.frame(date = Sys.Date()-7:1,
                                 max_length = sample(
                                   c(rep(0,20),0:13),
                                   replace =F,
                                   7
                                   )
)

avg_list_length <- mean(max_list_length_df$max_length[max_list_length_df$max_length!=0])
max_list_length <- max(max_list_length_df$max_length)
proportion_met_target <- 0.42

```

# Hello `r params$user_name`!

Thanks for submitting your recent `r taxon` records!

## Why record species 'lists'?

Recording as many species as possible provides scientists with valuable insights into the biodiversity of a given area. When we talk about species 'lists', we're referring to documenting more than just one or two striking or easily noticeable species—it's about observing and recording as many different species as you can find in a location.

```{r, results = 'asis'}
if(group >2){
  print(h2("Your recent lists"))
  
    
  print(p(paste0("Last week, you recorded an average of ",round(avg_list_length)," species per list. The longest list you submitted had ",max_list_length ," species.")))

  print(ggplot(max_list_length_df,aes(x = date,y = max_length))+
    geom_col(fill = "#92A41A")+
    theme_minimal(base_size = 18) +
    labs(x = "Date", y = "Longest list length")+
    scale_x_date(date_breaks = 'day', 
              date_labels = '%a')+
    geom_hline(yintercept = taxon_target,alpha= 1,linewidth = 2,colour = "red")+
    #annotate("text",x = Sys.Date()-4,y = 3,label = "List length target",colour = "red",size = 6)+
    lims(y = c(0,max(max_list_length,taxon_target+2))))
  

  
  if(max_list_length>=taxon_target){
    print(p(paste0("🥳 Well done, you achieved the challenge of submitting a list with ",taxon_target," or more species!")))
  } else {
    print(p(paste0("😢 Unfortunately, you didn’t achieve the challenge of submitting a list with at least ",taxon_target," species.")))
  }
}

if (group == 4){
  
  print(p(paste0(proportion_met_target*100,"% of recorders achieved the recent challenge!")))
}


```

## 🎯 Your challenge

This week, try to record a species list with at least `r taxon_target` species per list. Every extra species you record makes a big difference in our efforts to track biodiversity.

## Thanks!

Thank you for contributing to biodiversity monitoring! 

We’d love to hear your thoughts on this email! [Share your feedback here](`r paste0(params$config$controller_app_web_url,'submit_feedback/',params$content_key)`)

If you’d prefer to pause these updates, [unsubscribe here](`r paste0(params$config$controller_app_web_url,'unsubscribe/',params$content_key)`)
