---
title: "Recorder Feedback Example Email"
params:
    user_name: ""
    user_data: ""
    bg_data: ""
    bg_computed_objects: ""
    user_computed_objects: ""
    user_email: "UNKNOWN"
    content_key: ""
footer-date-time: "`r  format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`"
footer-unsubscribe-link: "www.google.com"
user-email: "`r params$user_email`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = FALSE, 
                      fig.align = 'center',
                      error=F,
                      results='asis')
```


```{r calculations, include=F}
#here you can do simple calculations for generating metrics that you later include in the text of the email. However if you wish to do more complex calculations it is recommended that these are done via prior # Get current date

current_date <- Sys.Date()
# Extract the month and year from the current date
current_month <- as.integer(format(current_date, "%m"))
current_year <- as.integer(format(current_date, "%Y"))
# Calculate the month and year of the previous month
previous_month <- current_month - 1
previous_year <- current_year
# Adjust for the previous year if the current month is January
if (previous_month == 0) {
  previous_month <- 12
  previous_year <- current_year - 1
}
# Create a date object for the first day of the previous month
first_day_of_previous_month <- as.Date(paste(previous_year, previous_month, "01", sep = "-"))
first_day_of_current_month <- as.Date(paste(current_year, current_month, "01", sep = "-"))
# Format the previous month
prev_month_string <- format(first_day_of_previous_month, "%B")

last_month_data <- params$user_data %>% 
  filter(date>=first_day_of_previous_month,date <first_day_of_current_month)

# a simple metric for number of records per day and a vector of species names
n_species <- last_month_data %>% pull(species) %>% unique() %>% length()
n_records <- last_month_data %>% nrow()
n_days <- last_month_data %>% 
  pull(date) %>% 
  unique() %>% 
  length()

species_names <- last_month_data %>% 
  group_by(species) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  pull(species)

records_per_day <- last_month_data %>% 
  group_by(date) %>% 
  summarise(number_of_records =n()) %>%
  mutate(date = as.Date(date))

this_year_data <- params$user_data %>% 
  filter(date>=as.Date(paste(current_year, "01", "01", sep = "-")),date<first_day_of_current_month)

n_species_year <- this_year_data %>% pull(species) %>% unique() %>% length()

sumulative_sp_count <- this_year_data %>% 
  group_by(species) %>% summarise(first_date = min(date))

cumulative_species_count <- this_year_data %>%
  arrange(date,species) %>%
  mutate(cumulative_species = cumsum(!duplicated(species))) %>%
  mutate(date = as.Date(date))



```

```{r packages, include =F}
library(dplyr) #wrangling
library(ggplot2) #plotting
library(ggspatial) #maps
library(sf)
library(htmltools)
```

```{css, echo=FALSE}
body{
  font-family: Helvetica;
  font-size: 13pt;
}

```

# Hello `r params$user_name`!

## Welcome to your iRecord month in review for `r prev_month_string`.

In `r prev_month_string` you recorded on `r n_days` days making a total of `r n_records` records. You recorded of `r n_species` different species. Your five most commonly recorded species were `r paste0(head(species_names,5),collapse = ", ")`.

```{r plot}
ggplot(records_per_day,aes(x = date,y = number_of_records))+
  geom_col(fill = "#92A41A")+
  theme_minimal(base_size = 18)+
  labs(x = "Date",y = "Number of records")
```

So far in `r current_year` you have recorded `r n_species_year` species.

```{r plot2}
ggplot(cumulative_species_count,aes(x = date,y = cumulative_species))+
  geom_step(colour = "#92A41A", size = 2)+
  theme_minimal(base_size = 18)+
  labs(x = "Date",y = "Number of species")
```

Thank you for your contribution!
